---
title: "hw1_401"
output: html_document
date: "2024-09-28"
---

```{r setup}
library(readr)
auto <- read.table("MLDS401/auto.txt", header = TRUE, stringsAsFactors = FALSE)
StockBeta <- read_csv("MLDS401/StockBeta.csv", show_col_types = FALSE)

auto$origin <- as.factor(auto$origin)
str(auto)
```

### problem 3

```{r problem_3_setup}
library(psych)
pairs.panels(auto[,c(2:8,1)], stars=T, density=T) # part a
round(cor(auto[,1:7], use="pair"),4)   # part b
fit = lm(mpg~., auto[,1:8])     # part c
summary(fit)

```

```{r problem_3_cdef}
# Assuming 'auto' is your dataset
library(psych)

auto$origin <-  factor((auto$origin), labels = c("US", "Europe", "Japan"))

# Fit the multiple linear regression model
fit = lm(mpg ~ cylinders + displacement + horsepower + weight + acceleration + year + origin, data = auto)

# Summary of the regression model
summary_fit = summary(fit)

# (c) F-test results
f_statistic = summary_fit$fstatistic
f_p_value = pf(f_statistic[1], f_statistic[2], f_statistic[3], lower.tail = FALSE)

cat("(c) F-statistic:", f_statistic[1], "\n")
cat("   p-value:", f_p_value, "\n\n")

# (d) Statistically significant predictors
cat("(d) Statistically significant predictors:\n")
significant_predictors = summary_fit$coefficients[summary_fit$coefficients[,4] < 0.05, ]
print(significant_predictors)
cat("\n")

# (e) and (f) Slope coefficients for year and displacement
year_coef = summary_fit$coefficients["year", ]
displacement_coef = summary_fit$coefficients["displacement", ]

cat("(e) Slope coefficient for year:", year_coef[1], "\n")
cat("   p-value:", year_coef[4], "\n\n")

cat("(f) Slope coefficient for displacement:", displacement_coef[1], "\n")
cat("   p-value:", displacement_coef[4], "\n")

```

a) Among all predictors, Cylinders, displacement, horsepower, and weight show relatively strong negative relationships with mpg. As these variables increase, mpg tends to decrease. While acceleration, year, and origin show relatively weak positive relationships with mpg. As these variables increase, mpg tends to increase as well.

b) The correlation between mpg and displacement is -0.80. This strong negative correlation indicates that as displacement increases, mpg tends to decrease significantly. Since all of the correlation coefficients are marked with three asterisks (***) between predictors and mpg, it indicates all variables have significant relationships with mpg.

c) Since the F-test is 224.4507, which is highly significant, it suggests that there is a statistically significant relationship between the set of predictors and the mpg response variable.

d) All predictors have p-values less than the significance level of 0.05, indicating that they all have a statistically significant relationship with the response variable mpg. Among these predictors, year has the most significant relationship since it has the lowest p-value and highest t-value.

e) The positive slope of 0.7770269 suggests that for each additional year (as cars get newer), the mpg increases by approximately 0.7770269 units, holding all other variables constant.

f) The positive slope of 0.02397864 suggests that for each unit increase in displacement, the mpg slightly increases by approximately 0.02397864 units, holding all other variables constant.


