---
title: "Final_project"
author: "BNarayanan"
date: "2024-11-24"
output: html_document
---
```{r}
library(dplyr)
file_path <- "np (1).csv"
np <- read.table(file_path, header = TRUE, sep = " ", na.strings = ".")
head(data)
```
```{r}
data <- data %>% select(-starts_with("Loc"), -starts_with("Src"))
str(data)
```
```{r}
colSums(is.na(data))
# For example, replace missing values with column means (numeric columns)
data <- data %>% mutate(across(where(is.numeric), ~ ifelse(is.na(.), mean(., na.rm = TRUE), .)))

```
```{r}
table(data$churn)

summary(select(data, regularity, intensity, currprice, trial))

```
```{r}
np <- np %>%
  arrange(SubscriptionId, t) %>% # Sort by customer ID and month
  group_by(SubscriptionId) %>%  # Group by each customer
  mutate(
    nextchurn = lead(churn),         # Churn status for next month
    nextprice = lead(currprice),     # Price paid for next month
    t = t                            # Retain the current time column
  )
head(np)
```
```{r}
# Model 1: nextchurn ~ t + trial + nextprice + regularity + intensity
model1 <- glm(nextchurn ~ t + trial + nextprice + regularity + intensity, 
              data = np, 
              family = binomial)

# Model 2: nextchurn ~ t + trial + nextprice + regularity
model2 <- glm(nextchurn ~ t + trial + nextprice + regularity, 
              data = np, 
              family = binomial)

# Model 3: nextchurn ~ t + trial + nextprice + intensity
model3 <- glm(nextchurn ~ t + trial + nextprice + intensity, 
              data = np, 
              family = binomial)

```

```{r}
# Summary for Model 1
summary(model1)

# Summary for Model 2
summary(model2)

# Summary for Model 3
summary(model3)
```

```{r}
cat("Model 1 AIC:", AIC(model1), "\n")
cat("Model 2 AIC:", AIC(model2), "\n")
cat("Model 3 AIC:", AIC(model3), "\n")
```

```{r}

library(car)
vif(model1)
vif(model2)
vif(model3)
```
### 2
##### Trial Effects
In all three models, trial has a positive coefficient and is statistically significant (p < 0.05). This suggests that customers on trial offers are more likely to churn.
This makes sense given the nature of trial offers, where customers may not transition into paying subscriptions after the trial period.
##### Price
In all models, nextprice has a positive coefficient and is statistically significant. This means that higher subscription prices are associated with an increased likelihood of churn.
Price sensitivity is a common driver of churn, suggesting that affordability or perceived value might be areas for improvement.
##### Regularity
Regularity has a negative coefficient and is highly significant in both Models 1 and 2. This indicates that more frequent engagement (measured by reading days per month) reduces the likelihood of churn.
The strong significance of regularity highlights the importance of creating habits and encouraging customers to engage with content consistently.

##### Intensity
Intensity has a negative coefficient and is statistically significant in Models 1 and 3. This suggests that deeper engagement (measured by page views per reading day) is also associated with lower churn.

However, the p-value in Model 1 (when both regularity and intensity are included) indicates that intensity is not as strong of a predictor compared to regularity.

##### VIFs
Diagnostics (VIFs):
The Variance Inflation Factors (VIFs) for all models are low (close to 1), indicating no significant multicollinearity among the predictors.

#### Part a
Short Trial Period:
Since most trial offers are for 1 month, the trial variable likely captures the behavior of customers during or immediately after their trial period.
A positive coefficient means customers on trials are more likely to churn, which could indicate that the trial experience is not compelling enough to retain them or that the transition to a paid subscription is a barrier (e.g., due to pricing or lack of perceived value).

Customers Without Trials:
Customers who are not on trial offers are less likely to churn, possibly because they’ve already committed to a paid subscription and see value in the product.
This highlights the need to improve the transition from trial to paid subscriptions, perhaps through personalized onboarding, discounted first payments, or additional benefits for trial users.

#### Part b
Regularity is more significant and has a larger effect (more negative coefficient) than intensity:
In Model 1, regularity is highly significant (p = 0.000176), while intensity is not significant (p = 0.135285).
In Models 2 and 3 (when analyzed independently), both are significant, but regularity has a larger effect size.

Focus on Regularity:
The organization should prioritize strategies that encourage customers to engage frequently. Examples include:
Daily content updates to give users a reason to return every day.
Push notifications or reminders to encourage regular visits.
Gamification features like streaks or rewards for consecutive days of use.

### 3
```{r}
model4 <- glm(nextchurn ~ t + trial + nextprice + sports1 + news1 + crime1 + 
                life1 + obits1 + business1 + opinion1, 
              data = np, 
              family = binomial)
summary(model4)
vif(model4)
cat("Model 4 AIC:", AIC(model4), "\n")
```
```{r}
model5 <- glm(nextchurn ~ t + trial + nextprice + sports1 + news1 + crime1 + 
                life1 + obits1 + business1 + opinion1 + regularity, 
              data = np, 
              family = binomial)

summary(model5)
cat("Model 5 AIC:", AIC(model5), "\n")
vif(model5)
```
The coefficient of regularity is -0.0288 and highly significant (p = 0.00137). This reinforces the finding that more frequent engagement reduces churn.
This result aligns with earlier models, showing that regularity plays a critical role in retaining customers.

Several content variables (e.g., sports1, news1, crime1) lose statistical significance after adding regularity:
sports1 becomes non-significant (p = 0.80243).
news1 is marginally non-significant (p = 0.12865).
Other content variables like crime1, life1, and obits1 remain non-significant.
This suggests that the effects of content engagement are largely explained by the customer's regularity (number of reading days).
trial, nextprice, and t remain significant with little change in their coefficients:
trial (p = 0.02935): Trial users are more likely to churn.
nextprice (p = 2.11e-06): Higher prices increase churn.
t (p = 1.02e-06): Longer subscription tenure reduces churn.

Including regularity in the model strengthens the conclusion that regular engagement is the most critical factor in reducing churn. 

### 4
```{r}
model6 <- glm(nextchurn ~ t + trial + nextprice + regularity + mobile + tablet + desktop, 
              data = np, 
              family = binomial)

# View the summary
summary(model6)
cat("Model 6 AIC:", AIC(model6), "\n")
vif(model6)
```
Mobile Usage:

Coefficient: 0.0017 (positive, very small magnitude).
Significance: p = 0.537 (not statistically significant).
Interpretation:
Mobile usage does not have a meaningful impact on churn in this model.
This suggests that the frequency of mobile sessions alone is not a strong determinant of whether a customer stays or churns.

Tablet Usage:

Coefficient: -0.0010 (negative, very small magnitude).
Significance: p = 0.821 (not statistically significant).
Interpretation:
Tablet usage has no significant impact on churn.
This could be due to tablet usage being more situational or infrequent, contributing less to overall engagement.

Desktop Usage:

Coefficient: -0.0026 (negative, very small magnitude).
Significance: p = 0.400 (not statistically significant).
Interpretation:
Desktop usage also shows no significant relationship with churn.
This could indicate that desktop sessions are not representative of engagement or retention behaviors.

Device Choice is Not a Key Driver of Churn:
Customers are equally likely to churn regardless of the devices they use to engage with the product.

Regularity remains significant and negatively associated with churn, reinforcing the idea that fostering consistent engagement (across any device) is more critical than focusing on specific platforms.

###5
```{r}
model7 <- glm(nextchurn ~ t + trial + nextprice + regularity +
                sports1 + news1 + crime1 + life1 + obits1 + business1 + opinion1 +
                mobile + tablet + desktop, 
              data = np, 
              family = binomial)

# View the summary
summary(model7)

# Calculate AIC
cat("Model 7 AIC:", AIC(model7), "\n")
vif(model7)
```
The inclusion of payment, content, and device variables does not substantially alter the conclusions:
Payment variables (trial, nextprice) and regularity remain key drivers of churn.
Content and device variables add little explanatory power, suggesting limited direct effects on churn
##### Lasso + Cross-Valid
```{r}
library(glmnet)
complete_data <- np[complete.cases(np[, c("nextchurn", "t", "trial", "nextprice", "regularity",
                                          "sports1", "news1", "crime1", "life1", "obits1",
                                          "business1", "opinion1", "mobile", "tablet", "desktop")]), ]
X <- model.matrix(nextchurn ~ t + trial + nextprice + regularity +
                    sports1 + news1 + crime1 + life1 + obits1 + business1 + opinion1 +
                    mobile + tablet + desktop, data = complete_data)[, -1]  # Exclude intercept

# Create response variable
y <- complete_data$nextchurn

# Perform cross-validated LASSO
cv_lasso <- cv.glmnet(X, y, family = "binomial", alpha = 1, nfolds = 10)

# Best lambda (penalty parameter)
best_lambda <- cv_lasso$lambda.min
cat("Best lambda:", best_lambda, "\n")

# Fit the final LASSO model using the best lambda
lasso_model <- glmnet(X, y, family = "binomial", alpha = 1, lambda = best_lambda)

# View coefficients
coef(lasso_model)
```
t: Negative coefficient (-0.1129) indicates that longer subscription tenure reduces churn.
trial: Positive coefficient (0.2861) shows that trial users are more likely to churn.
nextprice: Positive coefficient (0.0656) indicates that higher prices increase churn likelihood.
regularity: Negative coefficient (-0.0210) demonstrates that more frequent engagement reduces churn.
desktop: Slight negative coefficient (-0.0016), suggesting that higher desktop usage may slightly reduce churn.

All content variables (sports1, news1, crime1, etc.) were excluded.
Device variables like mobile and tablet were also excluded.

Only desktop is retained, with a negligible effect. This reinforces the conclusion that device type is not a key driver of churn.
